\begin{comment}
\documentclass[11pt]{article}
\usepackage[text={6.6in,9.3in},centering]{geometry}
\usepackage{times}
\usepackage{amsthm}
%\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amssymb, amsmath}
\usepackage{color}
\usepackage{graphicx}
%\usepackage{subfig}
\usepackage{lipsum}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{multicol}
\usepackage{algpseudocode}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{cleveref}
%\usepackage{algcompatible}

 \usepackage{pstricks}
% The following commands are not supported in PSTricks at present
% We define them conditionally, so when they are implemented,
% this pstricks file will use them.
\ifx\setlinejoinmode\undefined
  \newcommand{\setlinejoinmode}[1]{}
\fi
\ifx\setlinecaps\undefined
  \newcommand{\setlinecaps}[1]{}
\fi
% This way define your own fonts mapping (for example with ifthen)
\ifx\setfont\undefined
  \newcommand{\setfont}[2]{}
\fi

\algtext*{EndIf}% Remove "end if" text

%\newtheorem{theorem}{Theorem}[section]

\newtheorem{theorem}{Theorem}
\newtheorem{lemmas}{Lemma}
\newtheorem{axioms}{Axiom}
\newtheorem{lemma}[lemmas]{Lemma}
\newtheorem{axiom}[axioms]{Axiom}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{observation}[lemmas]{Corollary}
\newtheorem{invariant}{Invariant}

\newtheorem*{rep@theorem}{\rep@title}
\newcommand{\newreptheorem}[2]{%
\newenvironment{rep#1}[1]{%
 \def\rep@title{#2 \ref{##1}}%
 \begin{rep@theorem}}%
 {\end{rep@theorem}}}


\newreptheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
\newreptheorem{lemma}{Lemma}



\newcommand{\sq}{\hbox{\rlap{$\sqcap$}$\sqcup$}}
%\newcommand{\qed}{\hspace*{\fill}\sq}
%\newenvironment{proof}{\noindent {\bf Proof.}\ }{\qed\par\vskip %4mm\par}
\newenvironment{proofs}{\noindent {\bf Proof (sketch).}\ }{\qed\par\vskip
0mm\par}

\newcommand{\todo}[1]{\textcolor{red}{\textsf{[#1]}}}
\newcommand{\srb}[1]{\textcolor{blue}{\textsf{Steve: [#1]}}}
\newcommand{{\rh}}{{\widehat r}}
\newcommand{{\Rh}}{{\widehat R}}
\newcommand{\E}{{\mathcal E}}
\newcommand{\T}{{\mathcal T}}
\newcommand{\ww}{{\mathfrak{w}}}
\newcommand{\diam}{{\rm diam}}
\newcommand{\dist}{{\rm dist}}
\newcommand{\length}{{\rm length}}
\newcommand{\publish}{{\sl publish}}
\newcommand{\move}{{\sl move}}
\newcommand{\lookup}{{\sl lookup}}
\newcommand{\Read}{{\rm Read}}
\newcommand{\Write}{{\rm Write}}
\newcommand{\mysf}[1]{\textup{\sffamily #1}}
\newcommand{\CONGEST}{\mysf{CONGEST}}

\newcommand{\note}[1]{[[[{\bf #1}]]]}

\newcommand\outneighbors{\mathop{\mbox{$out$-$\mathit{neighbors}$}}}
\newcommand\inneighbors{\mathop{\mbox{$in$-$\mathit{neighbors}$}}}
\newcommand\maxweight{\mathop{\mbox{$max$-$\mathit{weight}$}}}
\newcommand\nonstrong{\mathop{\mbox{$non$-$\mathit{strong}$}}}


\begin{document}
\sloppy
\begin{titlepage}

\title{Distributed Garbage Collection for General Graphs}

\author{
Hari Krishnan\thanks{Corresponding author.} \thanks{Center for Computation and Technology, and Computer
Science and Engineering Division, Louisiana State University, LA 70803.}\\
{\em  Louisiana State University}\\
\url{hkrish4@tigers.lsu.edu}
\and
Steven R Brandt\thanks{Center for Computation and Technology, and Computer
Science and Engineering Division, Louisiana State University, LA 70803.} \\
{\em  Louisiana State University}\\
\url{sbrandt@cct.lsu.edu}
\and
Costas Busch\thanks{Computer Science and Engineering Division, Louisiana State
University, LA 70803.}\\
{\em Louisiana State University}\\
\url{busch@csc.lsu.edu}
\and
Gokarna Sharma\thanks{Department of Computer Science, Kent
State University, OH 44242}\\
{\em Kent State University}\\
\url{sharma@cs.kent.edu}
}

\date{}
\maketitle

\input{abstract}
\centerline{}
\centerline{}
\centerline{{\bf Regular Paper. Consider also for Brief Announcement.}}
\centerline{{\bf Eligible for best student paper award: Hari Krishnan is a full
time student.}}

\thispagestyle{empty}


\end{titlepage}
\end{comment}

\input{introduction}

\input{model}

\input{singlecollector}
%\input{figs/Diagram1}
\input{multicollector}

\input{conclusion}

\begin{comment}
\bibliographystyle{plain}
\bibliography{dgcbib}
\end{comment}
%\appendix
\input{appendix}


\begin{comment}
\end{document}
\end{comment}